What We Eval in the Shadows

### 2021-04-09 paper/main.tex
Patterns

The top-level environment in R is called the global environment. New
environments can be created using \c{new.env}. They can be provided a parent
environment which becomes the enclosing scope of the new environment.

@@

@@

\todo{Pierre: check the numbers - we want to know when it is an environment, when it is a list / data frame}
Most of the time, \ie in \packageEnvirEnvSitePercent of the sites, \eval is
passed an actual environment.

\c{NULL}, a \c{list} or a data frame, or an integer, happen in
\packageEnvirUnusualSitePercent of the cases: \eval copies the fields of the
list or data frame and creates bindings for them in a new environment.

@@

\mypara{Intercession.} \todo{Pierre: update wrt the previous section if we want
to keep it. Also perhaps we could start with a simpler pattern and then show
this one} A common use case for \eval is to be combined with \c{match.call}.
\c{match.call} walks up the call stack, captures the code that invoked the
currently executing function, and returns it as an unevaluated expression. The
pattern is to transform a call to some function \c{f} into a call to \c{g} with
some arguments retained and others modified.  \todo{or change the target
function. Both are described as "patterns" in the R documentation.} As an
illustration, consider the \c{vcpart} package's function \c{tvcglm} that is
translated to a call to \c{tvcm} with two modifications to the argument list:
argument \c{control} can't be missing and \c{fit} is set \c{"glm"}. The
function ends with a call to \c{eval.parent} to ensure that the rewritten call
is evaluted in the same environment the original call was.  \begin{lstlisting}
tvcglm <- function(formula, data, family, ...) { k <- match.call() k[[1]] <-
as.name("tvcm") if ("fit" %in% names(list(...))) warning("'fit' ignored.")
k$fit <- "glm" eval.parent(k) } \end{lstlisting} This pattern is recognizable
by the fact that the expression is a call and the target environment is that of
the parent.

@@

\mypara{Debloating.} \todo{Wouldn't reducing boilerplate be a better name?} \Eval is often used as a means to reduce boilerplate code;
simple and repetitive code can easily be replaced with judcious use of \eval.
For example, the \c{data.table} package uses \eval to calls the \c{options}
function with named arguments taken from a vector of strings. While the benefits
are limited in this example, it is an attractive tool for programmers.
\begin{lstlisting}
  opts = c("datatable.verbose"="FALSE", # ...many others
  for (i in names(opts))
    eval(parse(text=paste0("options(",i,"=",opts[i],")")))
\end{lstlisting}
This pattern is a special case of code generation, recognizable by the
fact that \eval is executed in a loop.

\mypara{Trivial.} When values are passed to \eval, they are returned
unchanged. They are an example of trivial uses of \eval. Another
trivial use is the empty expression, often found in JavaScript, but
rare in R. \todo{Given that we do not have any data I would get rid of it}

>>>>

The top-level environment in R is called the global environment. New
environments can be created using \c{new.env}, or \c{list2env}. They can be provided a parent
environment which becomes the enclosing scope of the new environment.
% as.environment also can be used sometimes for that
% list2env creates an environment only if its argument envir is not provided

@@

Most of the time, \ie in \packageEnvirEnvSitePercent of the sites, \eval is
passed an actual environment.

\c{NULL}, a \c{list} or a data frame, or an integer: \eval copies the fields of the
list or data frame and creates bindings for them in a new environment.

@@

@@

\mypara{Variable lookup.} % predominant usage of eval in base

\mypara{Symbolic differentiation.} % D and deriv

\mypara{Capturing a call.}  A common use case for \eval is to be combined with \c{match.call}.
\c{match.call} walks up the call stack, captures the code that invoked the
currently executing function, and returns it as an unevaluated expression that includes the call function and its unevaluated arguments. The
pattern is to transform a call to some function \c{f} into a call to \c{g} with
some arguments retained and others modified.  \todo{or change the target
function. Both are described as "patterns" in the R documentation.} As an
illustration, consider the \c{vcpart} package's function \c{tvcglm} that is
translated to a call to \c{tvcm} with two modifications to the argument list:
argument \c{control} can't be missing and \c{fit} is set \c{"glm"}. The
function ends with a call to \c{eval.parent} to ensure that the rewritten call
is evaluated in the same environment the original call was.  \begin{lstlisting}
tvcglm <- function(formula, data, family, ...) { k <- match.call() k[[1]] <-
as.name("tvcm") if ("fit" %in% names(list(...))) warning("'fit' ignored.")
k$fit <- "glm" eval.parent(k) } \end{lstlisting} This pattern is recognizable
by the fact that the expression is a call and the target environment is that of
the parent.

% We have quantitative data on that if needed...
This pattern is very often (more than half of the cases) used in relation with statistical models, to build a new model. \c{model.frame} to build a new generic model, or a more specialized model like \c{glm} or, \c{lm}.
% Show example

A related pattern is not to change the call but rather to evaluate the arguments in the parent frame. This can also be achieved with \c{substitute}.
% Example

In the R community, the call capturing pattern is often referred~\cite{hadley} as \emph{Non-Standard Evaluation}.

@@

\mypara{Debloating.} \todo{Wouldn't reducing boilerplate be a better name?} \Eval is often used as a means to reduce boilerplate code;
simple and repetitive code can easily be replaced with judicious use of \eval.
For example, the \c{data.table} package uses \eval to calls the \c{options}
function with named arguments taken from a vector of strings. While the benefits
are limited in this example, it is an attractive tool for programmers.
\begin{lstlisting}
  opts = c("datatable.verbose"="FALSE", # ...many others
  for (i in names(opts))
    eval(parse(text=paste0("options(",i,"=",opts[i],")")))
\end{lstlisting}
This pattern is a special case of code generation, recognizable by the
fact that \eval is executed in a loop.

\mypara{Trivial.} When values are passed to \eval, they are returned
unchanged. They are an example of trivial uses of \eval. Another
trivial use is the empty expression, often found in JavaScript, but
rare in R. \todo{Given that we do not have any data I would get rid of it}
\todo{We have the normalized expressions V and the small study I did on values passed to eval. And I did not find any eval() (empty) in the dataset. }

% Creating macros?
